Learning features and abstract actions for computing generalized plans

### 2018-09-05 doc/paper.tex
Some progress incorporating H's comments

% *Fill up general details in 1--6 that don't change from problem to problem
%how ${\cal S}$ obtained
The sample set $\mathcal{S}$ is obtained following one of two possible strategies.
In the \emph{exhaustive} sampling strategy,
$\mathcal{S}$ contains, for every one of the training instances $P$,
all the states generated on a breadth-first search from the initial state in $P$
that is run until $m$ expansions have been performed. If that set includes no goal state,
an optimal plan for $P$ is computed and all of the states along the plan are included in $\mathcal{S}$ as well,
as we need at least one goal state in the sample set.
%
In the \emph{random} sampling strategy, a large number $M$ of states are generated, potentially exhausting the full state space,
and then a certain number $m <\!< M$ of them is sampled uniformly at random. If no goal state is found among those $m$,
one random goal state from the larger set is drawn at random and added into the sample.

The feature complexity bound $k$ in the feature pool $\mathcal{F}$ is set to a value of $k=8$.
In our experiments, we increased this value to $k=16$ if no model was found
for the logical theories $T$ and $T_G$, and although it would be more effective
to devise an adaptive strategy to do so incrementally or through binary search,
that is beyond the scope of this work.
% theory $T({\cal S},{\cal F})$, in which case it is increased to $k=20$. 
% We have considered strategies to increase the bound adaptively
The resulting theories are then solved with the Open-WBO MAXSAT solver \cite{martins2014open}.

@@

The number of instances $P$ is never larger than $3$, and they are selected XXX THIS NEEDS TO BE WELL JUSTIFIED...

@@

\rowspacing \hline \rowspacing
\multirow{1}{*}{$\Q_{on\text{-}2\text{-}tow}$}   
& 3 &        420  & 8 & 657 & 128K & 25.8M & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - \\% Total complexity of F: 18
% Max-sat problem has 128217 variables and 25825428 unique clauses (with repetitions: 25825428). Top weight is 4577

@@

\paragraph{Clearing a block.}
The generalized problem $\Q_{clear}$ aims at achieving the goal $clear(x)$ for any block $x$
in the Blocks world. 
We take the standard encoding with primitive predicates $on(a, b)$, $clear(b)$, $ontable(b)$, $holding(b)$ and $handempty$.
%
With one single instance $\Q_{clear}$ with 5 blocks, we are able to learn
an abstract action model which features two actions:

@@

\paragraph{Stacking two blocks from different towers.}
The problem $\Q_{on\text{-}2\text{-}tow}$ is aims at achieving a goal of the form $on(x,y)$
in the Blocks world for any two blocks $x$ and $y$ that are initially not in the same tower.
We take the same encoding as above. Based on 3 different instances with up to 14
blocks, we are able to learn an abstract action model with 5 features and 7 actions.
The set of features is made up of  $X$, $n(x)$ and $n(y)$, with the same syntax and semantics as given before,
as well as $E: handempty$
and $XY: [(\exists on^{-1} . \set{x})\sqcap \set{b}]$,
the second of which precisely denotes that block $x$ is on block $y$.
The set of actions includes actions to pick up the topmost block above $x$, to pick up the topmost block above $y$,
to put a block being held \emph{aside}, i.e. not above $x$ nor above $y$, to pick $x$, 
to place $x$ on $y$ and to place $x$ on the table. This last action will indeed be necessary in case block $x$ is initially being held
but block $y$ is not clear.

@@

\paragraph{Gripper.}
The problem $\Q_{gripper}$ is based on the classical Gripper problem, where a robot with two grippers needs to 
move a number of blocks from one room into another \emph{target} room $x$, which we take to be a parameter of the problem,
each gripper of the robot being able to carry only one ball at a time.
We use the standard encoding with primitive predicates $at\text{-}robby(l)$, $at\text{-}ball(b, l)$, $free(g)$, $carry(b, g)$ that denote,
respectively, whether the robot (the ball) is at location $l$, whether gripper $g$ is free, and whether gripper $g$ carries ball $b$,
plus unary type predicates $room$, $ball$, and $gripper$.
%
We feed our pipeline with 2 training instances of the problem with only 2 rooms and at most 4 balls.
The resulting abstract action model is based on the following features:

@@

- \emph{leave-ball(s)-at-$x$}: \tuple{C>0, X;  C\downarrow, G \uparrow}

@@

- \emph{leave-$x$}: \tuple{X, C=0, G > 0;   \neg X}

@@

\paragraph{Collecting rewards in a maze.}
The objective in the problem $\Q_{reward}$ is to move in an $n \times m$ grid to pick up certain
rewards that are spread out in certain grid cells, while avoiding other cells which are considered as \emph{blocked}.
The encoding has primitive predicates $reward(c)$, $at(c)$, $blocked(c)$ and $adjacent(c_1, c_2)$ that denote
respectively the position of the rewards, of the agent, of the blocked cells, and the grid topology.
We use two single instances of the problem with size $5 \times 5$ and $4 \times 4$ and different distributions of blocks and rewards.
We learn a set with 2 features:

@@

\paragraph{Summary.}
XXX Perhaps mention general aspects of the results here, if space allows, e.g.:
\begin{itemize}
 \item Running times of the SAT solver are negligible.
 \item Minimization of precondition DNFs
 \item Theory T
 \item Sample sparsity - we use very few instances, rather few transitions.
 \item ** Mention: symbold DRL paper by Shanahan in 
\end{itemize}

>>>>

% *Fill up general details in 1--6 that don't change from problem to problem
%how ${\cal S}$ obtained
The sample set $\mathcal{S}$ is obtained following one of two possible strategies.
In the \emph{exhaustive} sampling strategy,
$\mathcal{S}$ contains, for every one of the training instances $P$,
all the states generated on a breadth-first search from the initial state in $P$
that is run until $m$ expansions have been performed. If that set includes no goal state,
an optimal plan for $P$ is computed and all of the states along the plan are included in $\mathcal{S}$ as well,
as we need at least one goal state in the sample set.
%
In the \emph{random} sampling strategy, a large number $M$ of states are generated, potentially exhausting the full state space,
and then a certain number $m \ll M$ of them is sampled uniformly at random. If no goal state is found among those $m$,
one random goal state from the larger set is drawn at random and added into the sample.

The complexity bound $k$ in the feature pool $\mathcal{F} = \mathcal{F}^k$ is set to a value of $k=8$,
with just one difference in the concept generation process: distance features are only generated for the last example,
meant to illustrate their usefulness.
% In our experiments, we increased this value to $k=16$ if no model was found
% for the logical theories $T$ and $T_G$, and although it would be more effective
% to devise an adaptive strategy to do so incrementally or through binary search,
% that is beyond the scope of this work.
% theory $T({\cal S},{\cal F})$, in which case it is increased to $k=20$. 
% We have considered strategies to increase the bound adaptively
The resulting theories are then solved with the Open-WBO MAXSAT solver \cite{martins2014open}.

@@

The number of instances $P$ used to generate the samples was kept small 
by choosing the instances $P$ manually, like a teacher. 
Very few instances turn out to be required for generating general plans.
The learning scheme is not supervised, as the training set are state 
transitions, not state action pairs to be generalized; but is not fully unsupervised either as currently 
implemented, with the instances $P$ from $\Q$ selected by hand.
We do not expect the results to be different if such instances
were to be sampled at random, but the number of instances required for 
producing sound abstract actions and valid general plans, as well as the size of the 
theories, would likely be higher.

All the results below are with goal marking using the theory $T_G({\cal S},{\cal F})$ that is
smaller than $T({\cal S,F})$ and produces smaller abstract action sets  $A_F$, which are forced
to be sound relative to the sample $\mathcal{S}$ but not complete, thus 
avoiding abstract actions that may make  planning harder and but are not relevant to the goal.
In general, the theory $T({\cal S},{\cal F})$ results in a significantly larger number of atoms and clauses,
as seen in \ref{tab:exp-result-data}, which makes the MAXSAT solving process much harder.

@@

\rowspacing \hline \rowspacing
\multirow{1}{*}{$\Q_{on}$}   
& 3 &        420  & 8 & 657 & 128K & 25.8M & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - \\% Total complexity of F: 18
% Max-sat problem has 128217 variables and 25825428 unique clauses (with repetitions: 25825428). Top weight is 4577

@@

\paragraph{Clearing a block.}
The generalized problem $\Q_{clear}$ aims at achieving the goal $clear(x)$ for any block $x$
in the Blocks world. 
We take the standard STRIPS representation with primitive predicates $on(a, b)$, $clear(b)$, $ontable(b)$, $holding(b)$ and $handempty$.
%
With one single instance $\Q_{clear}$ with 5 blocks, we are able to learn
an abstract action model which features two actions:

@@

\paragraph{Stacking two blocks from different towers.}
The problem $\Q_{on}$ is aims at achieving a goal of the form $on(x,y)$
in the Blocks world for any two blocks $x$ and $y$ that are initially not in the same tower.
We take the same STRIPS representation as above. Based on 3 different instances with up to 14
blocks, we are able to learn an abstract action model with 5 features and 7 actions.
The set of features is made up of  $X$, $n(x)$ and $n(y)$, with the same syntax and semantics as given before,
as well as $E: handempty$
and $XY: [(\exists on^{-1} . \set{x})\sqcap \set{b}]$,
the second of which precisely denotes that block $x$ is on block $y$.

The set of actions includes actions to pick up the topmost block above $x$, to pick up the topmost block above $y$,
to put a block being held \emph{aside}, i.e. not above $x$ nor above $y$, to pick $x$, 
to place $x$ on $y$ and to place $x$ on the table.

- \emph{pick-up-above-$x$}:

- \emph{pick-up-above-$y$}:

This last action will indeed be necessary in case block $x$ is initially being held
but block $y$ is not clear.

@@

\paragraph{Gripper.}
The problem $\Q_{gripper}$ is based on the classical Gripper problem, where a robot with two grippers needs to 
move a number of blocks from one room into another \emph{target} room $x$, which we take to be a parameter of the problem,
each gripper of the robot being able to carry only one ball at a time.
We use the standard STRIPS representation with primitive predicates $at\text{-}robby(l)$, $at\text{-}ball(b, l)$, $free(g)$, $carry(b, g)$ that denote,
respectively, whether the robot (the ball) is at location $l$, whether gripper $g$ is free, and whether gripper $g$ carries ball $b$,
plus unary type predicates $room$, $ball$, and $gripper$.
%
We feed our pipeline with 2 training instances of the problem with only 2 rooms and at most 4 balls.
The resulting abstract action model is based on the following features:

@@

- \emph{drop-ball(s)-at-$x$}: \tuple{C>0, X;  C\downarrow, G \uparrow}

@@

- \emph{drop-$x$}: \tuple{X, C=0, G > 0;   \neg X}

@@

\paragraph{Collecting rewards in a maze.}
The objective in the problem $\Q_{reward}$ is to move in an $n \times m$ grid to pick up certain
rewards that are spread out in certain grid cells, while avoiding other cells which are considered as \emph{blocked}.
This example is from recent work in symbolic deep
reinforcement learning, where there is a penalty associated for 
stepping in blocked cells instead \cite{garnelo2016towards}.
%
We represent the instances $P$ with primitive predicates $reward(c)$, $at(c)$, $blocked(c)$ and $adjacent(c_1, c_2)$ that denote
respectively the position of the rewards, of the agent, of the blocked cells, and the grid topology.
We use two single instances of the problem with size $5 \times 5$ and $4 \times 4$ and different distributions of blocks and rewards.
We learn a set with 2 features:

@@

\paragraph{Observations.}
We conclude the section with a few general observations:

\begin{enumerate}
\item
\emph{Soundness}: interestingly, abstract actions obtained from a small number of samples in
$\mathcal{S}$ turn out to be sound not just for $\mathcal{S}$ but for the generalized problem $\Q$,
as discussed above. Yet this does not need to be true for the resulting policies to be 
correct for $\Q$.
%
Abstract actions $A_F$ do not actually have to be sound for a policy $\pi$
obtained from them to be sound in $\Q$.
Indeed, for the correctness of $\pi$ for solving $\Q$,
it suffices for $A_F$ to be sound \emph{over the states reachable by $\pi$}
in all the instances $P$ of $\Q$.

For example, the abstract action \emph{pick-up-above-$y$} in $\Q_{on}$
is sound relative to the sample set $\mathcal{S}$ used, but not relative to the general
$\Q_{on}$. In any state $s$ where $x$ is a few blocks above $y$ and clear,
there is no concrete action that instantiates \emph{pick-up-above-$y$},
as picking up the topmost block above $y$ would necessarily make feature $X \equiv holding(x)$
true, which the abstract action does not.
%
Keeping track of the information necessary to ensure general correctness would require indeed 
a feature set $F$ with higher complexity.
Yet if we start in states where $x$ and $y$ are in different towers, 
states as $s$ will never be reached, and hence the policy is correct and solves $\Q_{on}$.

\item
\emph{Grammar and Feature Pool}: the general DL grammar we have described appears to be 
adequate to generate a sufficiently expressive pool of features for the examples we use,
but this does not have to be always so.
To come up with a general strategy in the Towers of Hanoi, for example,
the first action to be taken  (whether to move the smallest disk to target or buffer peg) 
depends on the number of disks being odd or even,
which cannot be easily expressed with the type of features generated by our grammar.

\item
\emph{Expressive and Computational Limitations}:
The current approach

*** XXX ***

The use of primitive \emph{goal predicates} like $on_g$ with denotation
given by the goal conjunctions in the concrete instances, as done in 
\cite{martin-geffner:generalized}, could potentially alleviate this
e.g. to be able to generate general policies for all BW problems.

Why pipeline cannot be used to solve arbitrary problems. Eg, logistics.
Que partes explotan? Sizes k and $\abs{\mathcal{S}}$?

* Goal in $\Q$ needs to be expressible in generic terms (but this is not well formalized in our formulation)
* Increasing number of domain parameters kills scalability.
* How to select which are the good parameters if we want arbitrary goals? This is also related to the proper formalization
  of the approach, which we havent done.

%  \item Running times of the SAT solver are negligible.
%  \item Minimization of precondition DNFs
%  \item Theory T
\end{enumerate}
